RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-25 discussion.tex
merge
working

@@

% limitations
Previous studies have pointed out difficulties in achieving the flexibility,
scalability and maintainability expected of production software in experimental
code written by data analysts.
Although these properties cannot be enforced by any programming environment,
suitable tools can help well-motivated analysts and programmers to create
robust applications with much less effort.

>>>>

What are the consequences of introducing this technology into the
data analysis ecosystem.

Good and bad.

We have observed that people start sharing ideas through the system,
instead of hacking up prototypes and sending screenshots and Powerpoint
they hack up experiments and send them. Makes this easier to collaborate
about code but also makes it easy to collaborate about services that are
deployed.

The negative is that at every stage everything is deployed, so that if
you break something

Allows people to fork and try something else without restriction.
Have not solved the problem of having a lot of bits and pieces lying around
but have made it easier to get at them and to use algorithms to 
help organize them.

Recommendations - was done in Vistrails - within the same graph but
there is clustering of workflows that could be applied to recommending
across graphs.

Because we developed the ideas and the system together

Package dependencies and versions could be a problem

@@

% limitations
Previous studies have pointed out difficulties in achieving the flexibility,
scalability and maintainability expected of production software,
with experimental code written by data analysts. In RCloud this problem
is compounded by its flexible versioning, which may make it even
more challenging to ensure that production services are stable.
Versioning happens in several places: in scripts (that are under
our control in RCloud through git), in libraries (that should be
under the control of the R environment, though support for this
is very weak), and in the external environment
(such as operating system components and even the protocols
spoken by external services, for which there is no realistic
hope of formal control).  
Also, because there is no explicit separation between the experimental
and production environments in RCloud, it seems relatively easy for
data scientists to inintentionally modify or break production services.

Upgrading is troublesome, but what is better. When different things are being
upgraded in flight how do you ensure the things you are calling behave as
expected, and if someone upgrades 
Would require support from the R environment which does not exist now.

have only the right version of the script we are working on but
can't do the same for the environment
the environment includes remote services not under our control

Although these properties cannot be enforced by any programming environment,
suitable tools can help well-motivated analysts and programmers to create
robust applications with much less effort.
